---
title: "Firearm Homicide"
author: "Ohio Alliance for Population Health"
format:
  html:
    theme: united
    font:
    fontsize: 1.0em
output:
  html_document: default
  pdf_document: default
editor_options:
  chunk_output_type: console
self-contained: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = FALSE, echo = FALSE, warning = FALSE, message = FALSE)
```

## Background 
The OHIO Alliance for Population Health (OAPH) is a statewide collaborative focused on improving the health of all Ohioans by solving the most complex and pressing population health concerns across the state. In line with OAPH’s overarching mission, this report aims to identify disparities in gun violence mortality among Ohioans and consider how these disparities can be addressed to improve health and wellbeing across the state. 

## Methodology 
This report examines firearm homicide deaths across national and state levels over an extended period of time.  
 
CDC WONDER collects and maintains data regarding deaths and their causes. All data were extracted from CDC WONDER using provisional or final mortality data.  
 
All maps, graphs, and other figures for the United States and Ohio that are presented in this report were produced using R, RStudio, ggplot2, sf, tidyverse, and tidycensus. 

## Key Findings 
* In 2022, firearms accounted for 79.1% of all homicide deaths in the United States 
* In 2022, firearms accounted for 79.4% of all homicide deaths in Ohio 
* In the past 10 years (2013-2022), homicide deaths by firearms have increased 66.9% 
* Since 1999, homicide deaths by firearm have increased 52.5%  
* In 2022 Ohio ranked 32 (out of 51 states, with rank 1 indicating the lowest rate) in firearm homicide mortality with a reported rate of 6.3 deaths per 100,000 population, 6.8% higher than the national average 
* Americans ages 20-24 experience the highest rates of gun violence mortality  
* Ohioans ages 20-24 experience the highest rates of gun violence mortality  
* In 2022, men were over five times as likely to be killed by gun violence than women 

```{r loading packages, eval=FALSE}
library(tidyverse)
library(stringr)
library(scales)
library(zoo)
library(sjmisc)
library(janitor)
library(dplyr)
library(tidycensus)
library(tigris)
library(sf)
library(dplyr)
library(classInt)
library(extrafont)
```

```{r census geometry data}
#census_api_key("e91d231563691e6bea782a7dbafe82616924d422", overwrite = TRUE, install = TRUE)
#census_api_key('e91d231563691e6bea782a7dbafe82616924d422')

state <- get_acs(
  geography = "state",
  variables = "B01001_001",
  year = 2019,
  survey = "acs1",
  geometry = TRUE,
  resolution = "20m"
) %>%
  shift_geometry()

county <- get_acs(
  geography = "county",
  variables = c("B01001_001"),
  year = 2019,
  geometry = TRUE,
)

state %>% mutate(GEOID=as.integer(GEOID)) -> state
state %>% filter(NAME!="Puerto Rico") -> state
county %>% mutate(GEOID=as.integer(GEOID)) -> county
county %>% filter(NAME!="Puerto Rico") -> county

```

## Firearm Homicide Deaths by State, 2022

```{r 2022 national map by state}

read.delim(
  here::here(
    "data",
    "US.AllFirearmHomicide.State.2022.txt"
    )
  ) -> homicide

#Create calculated crude rate variable
homicide %>% mutate(Rate=(Deaths/Population)*100000) -> homicide

#Clean up data
homicide %>% filter(Notes=='') -> homicide
homicide = subset(homicide,select=-c(Notes)) 
homicide %>% rename(GEOID=Residence.State.Code) -> homicide

left_join(state,homicide, by='GEOID') -> homicide

#Create data breaks
breaks_j <- classIntervals(c(min(homicide$Rate),       
                             homicide$Rate),n=3,        
                           style="quantile")              

homicide <- mutate(homicide,Breaks=cut(Rate,breaks_j$brks,include.lowest=TRUE))

#Change legend category labels based on breaks
c('18.4 - 27.4','27.5 - 34.0','34.1 - 59.2') -> legend.labels

#Create map
ggplot() +
  
  #ggtitle("Firearm Homicide Deaths by State, 2022") +
  
  geom_sf(data = homicide, aes(fill = Breaks), color = "black",size=.4, alpha=1) +
  
  scale_fill_manual(values=c("#BFCEDB",
                             
                             "#6084A6",
                             
                             "#003A70"), na.value = "white", drop = FALSE, name="Rate per 100,000 population", label=legend.labels) +
  
  
  #geom_sf_text(data=state,aes(label= Code), color = "black", size = 2.5, vjust = -.5) +
  #geom_sf_text(data=State,aes(label= round(Rate,2)), color = "black", size = 2.5, vjust = 1.2) +
  
  theme_void() +
  
  theme(text = element_text(),
        plot.title = element_text(size = 20, color = '#004023', hjust = .5, face='bold'),
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) +
  theme(legend.position = 'bottom')
```

## Firearm Homicide Deaths by State, 2022

```{r 2022 state horizontal bars}

read.delim(
  here::here(
    "data",
    "US.AllFirearmHomicide.State.2022.txt"
    )
  ) -> homicide

homicide %>% mutate(Rate=(Deaths/Population)*100000) -> homicide

homicide %>% filter(Notes=='') -> homicide
homicide = subset(homicide,select=-c(Notes)) 
homicide %>% rename(State=Residence.State) -> homicide

#Create rank variable based on rate
homicide %>% mutate(Rank = dense_rank(Rate)) -> homicide

#Sort by high to low rate
homicide$State <- factor(homicide$State) %>% 
  fct_reorder(homicide$Rate)

#Create horizontal bar chart
ggplot(homicide,aes(x=State, 
                    y=Rate, 
                    fill=Rate)) + 
  geom_col(alpha=0.75) +   	#Alpha controls density of color, 
  
  #ggtitle("homicide Deaths by State, 2022") +
  
  #Creates breaks in graph & controls limits(min/max)
  #scale_y_continuous(limits = c(-0.2, 0.7), breaks = c(-0.2,-0.1,0,0.1,0.2,0.3,0.4,0.5,0.6,0.7)) +
  
  #Flips graph from vertical bars to horizontal
  coord_flip() +
  labs(caption = "", 
       y="Rate per 100,000 population",    		#Legend titles
       x="") +
  
  #Create fill gradient in bars
  scale_fill_gradient(name="Label",
                      low = "#BFCEDB",
                      high = "#003A70") + 	
  
  #Create average line
  geom_hline(yintercept = mean(homicide$Rate), color="grey40", linewidth=1, alpha=.5) +
  annotate('text', x=5, y=5.8, color='grey40', fontface='bold', label="National Average") +
  
  #Control rate numbers on graph
  geom_text(aes(label=round(Rate,1)), color='black', size=3, fontface="bold", hjust=1.2, vjust=.35) +
  
  #Control graph theme
  theme_minimal() +
  theme(text = element_text()) +
  theme(plot.title=element_text(size = 38, color = '#004023', hjust = 0, face='bold')) +
  theme(axis.text.x = element_text(color="gray20", 
                                   size=10),
        axis.text.y = element_text(color="black", 
                                   size=10)) +
  theme(legend.position = 'none')

```


## Firearm Homicide Deaths by State, 2018-2022

```{r 2018-22 national map by state}

read.delim(
  here::here(
    "data",
    "US.AllFirearmHomicide.State.2018-22.txt"
    )
  ) -> homicide

homicide %>% mutate(Rate=(Deaths/Population)*100000) -> homicide

homicide %>% filter(Notes=='') -> homicide
homicide = subset(homicide,select=-c(Notes)) 
homicide %>% rename(GEOID=Residence.State.Code) -> homicide

left_join(state,homicide, by='GEOID') -> homicide

#Create data breaks
breaks_j <- classIntervals(c(min(homicide$Rate),       
                             homicide$Rate),n=3,        
                           style="quantile")              

homicide <- mutate(homicide,Breaks=cut(Rate,breaks_j$brks,include.lowest=TRUE))

#Change legend category labels based on breaks
c('0.9 - 2.6','2.7 - 6.0','6.1 - 20.5') -> legend.labels

#Create map
ggplot() +
  
  #ggtitle("Firearm Homicide Deaths by State, 2018-2022") +
  
  geom_sf(data = homicide, aes(fill = Breaks), color = "black",size=.4, alpha=1) +
  
  scale_fill_manual(values=c("#BFCEDB",
                             
                             "#6084A6",
                             
                             "#003A70"), na.value = "white", drop = FALSE, name="Rate per 100,000 population", label=legend.labels) +
  
  #geom_sf_text(data=State,aes(label= Code), color = "black", size = 2.5, vjust = -.5) +
  #geom_sf_text(data=State,aes(label= round(Rate,2)), color = "black", size = 2.5, vjust = 1.2) +
  
  theme_void() +
  
  theme(text = element_text(),
        plot.title = element_text(size = 19, color = '#004023', hjust = .5, face='bold'),
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) +
  theme(legend.position = 'bottom')
```

## Firearm Homicide Deaths by State, 2013-2017

```{r 2013-17 national map by state}

read.delim(
  here::here(
    "data",
    "US.AllFirearmHomicide.State.2013-17.txt"
    )
  ) -> homicide

homicide %>% mutate(Rate=(Deaths/Population)*100000) -> homicide

homicide %>% filter(Notes=='') -> homicide
homicide = subset(homicide,select=-c(Notes)) 
homicide %>% rename(GEOID=State.Code) -> homicide

left_join(state,homicide, by='GEOID') -> homicide

#Create data breaks
breaks_j <- classIntervals(c(min(homicide$Rate),       
                             homicide$Rate),n=3,        
                           style="quantile")              

homicide <- mutate(homicide,Breaks=cut(Rate,breaks_j$brks,include.lowest=TRUE))

#Change legend category labels based on breaks
c('0.6 - 1.9','2.0 - 4.4','4.5 - 12.6') -> legend.labels

#Create map
ggplot() +
  
  #ggtitle("Firearm Homicide Deaths by State, 2013-2017") +
  
  geom_sf(data = homicide, aes(fill = Breaks), color = "black",size=.4, alpha=1) +
  
  scale_fill_manual(values=c("#BFCEDB",
                             
                             "#6084A6",
                             
                             "#003A70"), na.value = "white", drop = FALSE, name="Rate per 100,000 population", label=legend.labels) +
  
  #geom_sf_text(data=State,aes(label= Code), color = "black", size = 2.5, vjust = -.5) +
  #geom_sf_text(data=State,aes(label= round(Rate,2)), color = "black", size = 2.5, vjust = 1.2) +
  
  theme_void() +
  
  theme(text = element_text(),
        plot.title = element_text(size = 19, color = '#004023', hjust = .5, face='bold'),
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) +
  theme(legend.position = 'bottom')
```

## U.S. Firearm Homicide Deaths by Year, 1999 - 2022

```{r us time series line chart 1999-2022}

read.delim(
  here::here(
    "data",
    "US.AllFirearmHomicide.Year.1999-2020.txt"
    )
  ) -> homicide

read.delim(
  here::here(
    "data",
    "US.AllFirearmHomicide.Year.2021-22.txt"
    )
  ) -> homicide2

#Combine datasets
rbind(homicide,homicide2) -> homicide
homicide = subset(homicide,select=-c(Notes)) 

#Create calculated crude rate variable
homicide %>% mutate(Rate=(Deaths/Population)*100000) -> homicide

#Create plot
ggplot(homicide, aes(x=Year.Code,y=Rate)) +
  
  #ggtitle("U.S. Firearm Homicide Deaths by Year \n 1999 - 2022") +
  
  #Show trend line (smoothed data)
  geom_smooth(color="#004023",span=0.75, 
  se=TRUE, linewidth=1,fill="gray85") +  
  
  geom_col(fill="#003a70",alpha=.80) + 
  
  geom_text(aes(label= round(Rate, 1)), size=3.3, color ="black", fontface='bold', vjust=1.75) +
  
  
  labs(caption='', y="Rate per 100,000 Population", x=NULL, size=15) +
  
  scale_y_continuous(labels = comma) +
  #  limits = c(MIN, MAX).
  
  scale_x_continuous(limits = c(1998, 2023), 
                     breaks = c(1999,2001,2003,2005,2007,2009,2011,2013,2015,2017,2019,2021)) +
  
  #One theme (title left, only vertical background lines)  
  #theme_minimal()+theme(axis.title.y=element_text(size=10,colour="gray40")) +
  #theme(axis.title.x=element_text(size=10,colour="gray40")) +
  #theme(axis.ticks.x = element_line(size = .03), panel.grid.minor = element_blank(),
          #     panel.grid.major.y = element_line( size=.15, color="gray10", linetype = 'dotdash'),
          #   panel.grid.major.x = element_blank(),
   #axis.ticks.length=unit(1,"cm")) +
  #theme(legend.position = 'none')
  
theme_minimal() +
  theme(text = element_text()) +
  theme(plot.title=element_text(size = 30, color = '#003A70', hjust = 0, face='bold')) +
  theme(axis.text.x = element_text(color="black", size=11, vjust=8),
        axis.text.y = element_text(color="black", size=11, hjust=2)) +
  theme(panel.grid.major.y = element_line( size=.3, color="gray40", linetype = 'solid'),
        panel.grid.minor.y = element_line( size=.3, color="gray80", linetype = 'solid'),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()) +
  theme(legend.position = 'none')
```

## Ohio Firearm Homicide Deaths by County, 1999-2020

```{r oh map by county 1999-2020}

read.delim(
  here::here(
    "data",
    "OH.AllFirearmHomicide.County.1999-2020.txt"
    )
  ) -> homicide

#Create calculated crude rate variable
homicide %>% mutate(Deaths=as.integer(Deaths)) -> homicide
homicide %>% mutate(Rate=(Deaths/Population)*100000) -> homicide

#Clean up data
homicide %>% filter(Notes=='') -> homicide
homicide = subset(homicide,select=-c(Notes)) 
homicide %>% rename(GEOID=County.Code) -> homicide

left_join(county,homicide, by='GEOID') -> homicide
homicide %>% filter(County != 'NA') -> homicide
homicide %>% mutate(Name=str_sub(County,1,-12)) -> homicide

#Create data breaks
breaks_j <- classIntervals(c(min(homicide$Rate),       
                             homicide$Rate),n=3,        
                           style="quantile")              

homicide <- mutate(homicide,Breaks=cut(Rate,breaks_j$brks,include.lowest=TRUE))

#Change legend category labels based on breaks
c('0.5 - 1.1','1.2 - 2.2','2.3 - 9.1') -> legend.labels

#Create map
ggplot() +
  
  #ggtitle("Ohio Firearm Homicide Deaths by County, 1999-2020") +
  
  geom_sf(data = homicide, aes(fill = Breaks), color = "black",size=.4, alpha=1) +
  
  scale_fill_manual(values=c("#BFCEDB",
                             
                             "#6084A6",
                             
                             "#003A70"), na.value = "white", drop = FALSE, name="Rate per 100,000 population", label=legend.labels) +
  
  
  geom_sf_text(data=homicide,aes(label= Name), size = 2.5, vjust=-.25,
               color = case_when(homicide$Breaks == '[0.467,1.07]' | homicide$Breaks == '(1.07,2.22]' | homicide$Rate == '' ~ 'black',
                                 homicide$Breaks == '(2.22,9.13]' ~ 'white')) +
  #geom_sf_text(data=State,aes(label= round(Rate,2)), color = "black", size = 2.5, vjust = 1.2) +
  
  theme_void() +
  
  theme(text = element_text(),
        plot.title = element_text(size = 38, color = '#004023', hjust = .5, face='bold'),
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) +
  theme(legend.position = 'bottom')

```

